---
title: "Who Counts as Australian?"
author:
  - Kelly Tall
  - Michael Falk
---

```{r message=FALSE, include=FALSE}
devtools::install_github("wikihistories/wikkitidy")
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE
)
library(tidyverse)
library(igraph)
library(wikkitidy)
data_dir <- "data/who-counts"
```

Who counts as Australian in Wikipedia and its sister-projects? How does Wikipedia portray Australia's population? These questions are easy to ask but hard to answer, because of Wikipedia's *representational multiplicity*. Wikipedia represents Australia in many ways. Its sister projects—Wikidata and Wikimedia Commons preminent among them—represent Australia in myriad other ways.

In this report, we consider several ways that Australians are represented in Wikipedia and Wikidata. We focus on four main modes of representation:

- [Textual representation]: How Wikipedia answers the question 'who counts as Australian' in its articles on the topic
- [The category system]: How Wikipedia categorises people as 'Australian'—including its categorisation of Indigenous Australians
- [The search interface]: How Wikipedia selects and sorts 'Australians' when you search the encyclopaedia
- [Wikidata properties]: How Wikidata splits the concept of 'Australian' across different properties in the database

In [Conclusion: beyond bias], we make some suggestions about how Wikipedians might like to improve the representation of Australians in the encyclopaedia. As well as addressing the traditional problem of 'systemic bias', we suggest that Wikipedians draw attention to the *representational multiplicity* of Wikipedia, so that visitors to the site can become more aware of the multiple definitions of 'Australianness' across the Wikiverse. 

## Textual representation

The first answer Wikipedia gives to the question 'Who counts as Australian?' is: [any 'individual' with a 'residential, legal, historical or ethno-cultural' connection to 'the country of Australia'](https://en.wikipedia.org/w/index.php?title=Australians&oldid=1150975782).[^1] At the time of writing, the 'Australians' comprise an estimated [26,468,200](https://en.wikipedia.org/w/index.php?title=Australia&oldid=1153684876) residents of Australia, and a diaspora of [577,255](https://en.wikipedia.org/w/index.php?title=Australians&oldid=1150975782) Australians overseas.

[^1]: Throughout the report, we link to the particular revision of any page that we consulted in our research.

TODO: Do we need this section? Where should we go with it? 

## The category system

```{r}
get_aussies_category_tree <- function(use_cache = T) {
  data_files <- list.files(data_dir)
  aussie_cat_files <- stringr::str_detect(data_files, "australian-people-categories")
  if (!use_cache || sum(aussie_cat_files) == 0) {
    # This takes a long time (> 1 hour)
    australians <- wikkitidy::build_category_tree("Category:Australian people")
    readr::write_rds(
      australians,
      file.path(
        data_dir,
        glue::glue("australian-people-categories-{as.integer(Sys.time())}.rds")
        ),
      compress = "gz"
    )
  } else {
    latest_file <- sort(data_files[aussie_cat_files], decreasing = TRUE)[1]
    australians <- readr::read_rds(file = file.path(data_dir, latest_file))
  }
  australians
}
australians <- get_aussies_category_tree()
```

The second answer Wikipedia gives to the question, 'Who counts as Australian?' is: everyone who falls under [Category:Australian people](https://en.wikipedia.org/wiki/Category:Australian_people). There are nearly 100,000 articles on English Wikipedia that relate to Australian people, and the number is growing consistently (@fig-included-in-ozpeople-over-time).[^2]

[^2]: The growth was calculated using the 'timestamp' data returned by [API:Categorymembers](https://www.mediawiki.org/wiki/API:Categorymembers). The 'timestamp' records when an article, file or subcategory was included in a given category. The documentation does not explain how the inclusion time is calculated for pages that are removed from and then re-included in a category. If an article was included in multiple subcategories of Category:Australian people, then the earliest timestamp was used to calculate its position in the figure.

```{r}
#| label: fig-included-in-ozpeople-over-time
#| fig-cap: Category:Australian people has grown consistently since the early days of Wikipedia
australians$nodes %>% 
  filter(ns == 0) %>%
  left_join(australians$edges, by = join_by(pageid == target)) %>% 
  mutate(included = as_datetime(timestamp)) %>% 
  arrange(pageid, included) %>% 
  group_by(pageid) %>% 
  slice(1) %>% 
  ungroup() %>% 
  mutate(included = round_date(included, "day")) %>% 
  count(included) %>% 
  ggplot(aes(included, cumsum(n))) +
  geom_line() +
  labs(x = "Time (days)", y = "Cumulative number of articles")
```
```{r}
category_graph <- graph_from_data_frame(
  d = australians$edges %>% 
    left_join(australians$nodes, join_by(target == pageid)) %>% 
    filter(ns == 14) %>% 
    select(source:timestamp),
  directed = TRUE,
  vertices = australians$nodes %>% filter(ns == 14)
)

vertex_by_id <- function(graph, pageid) {
  V(graph)[V(graph)$name == pageid]
}

root <- V(category_graph)[[1]]

category_url <- function(category, collapse = " → ") {
  paste0("[", category, "](https://en.wikipedia.org/wiki/", category, ")", collapse = collapse)
}
```

When editors include articles in [Category:Australian people](https://en.wikipedia.org/wiki/Category:Australian_people), they do not necessarily do so intentionally. Category:Australian people is the root node of a vast tree-like [graph](https://en.wikipedia.org/wiki/Graph_(discrete_mathematics)) of categories. The editor who categorised [Steve Wozniak](https://en.wikipedia.org/wiki/Category:Australian_people) as a [staff member of the University of Technology Sydney](https://en.wikipedia.org/wiki/Category:Academic_staff_of_the_University_of_Technology_Sydney) was probably unaware that they were thereby categorising Wozniak as an 'Australian person'. The category for academic staff at UTS lies several layers deep in the category graph, `r distances(category_graph, root, vertex_by_id(category_graph, 32435264))` steps away from the root: `r shortest_paths(category_graph, root, vertex_by_id(category_graph, 32435264))$vpath[[1]]$title %>% rev() %>% category_url()`. Perhaps an editor may quibble that a person who works at UTS is not necessarily 'from Sydney', but in the present state of the category graph, there is no other way to mark a staff member of UTS. The category graph is strictly hierarchical. A subcategory cannot be removed from its parent category for the sake of a single page.

```{r}
full_graph <- graph_from_data_frame(
  d = australians$edges,
  directed = T,
  vertices = australians$nodes
)

# These are the distances from the *article* to the root category. To get the
# weighted distance of the subcategories to the root, take mean(article_distances)
# and subtract one.
article_distances <- distances(
    full_graph,
    v = V(full_graph)[[1]], 
    to = V(full_graph)[V(full_graph)$name %in% filter(australians$nodes, ns == 0)$pageid]
    )
```

Category:Australian people has `r length(category_graph) %>% prettyNum(",")` subcategories, many of which are even farther from the root than the staff of UTS. We can get a sense of how vast this structure is by taking [diameter](https://en.wikipedia.org/wiki/Distance_(graph_theory)#Related_concepts) of the category graph. In this case, the diameter is `r diameter(category_graph)`, which means that the most distant subcategory of Category:Australian people is `r diameter(category_graph)` steps away. Such distant nodes are called 'eccentric' in graph theory. In this case, the most distant subcategory truly is eccentric: `r farthest_vertices(category_graph)$vertices$title[[2]] %>% category_url()` does not include articles about 'Australian people' at all, but about indie rock records. This case is extreme. The case of UTS staffmembers is more typical. If we consider all the articles that have been placed in Category:Australian people or its subcategories, we find that the subcategories used are on average `r (mean(article_distances) - 1) %>% round(2)` steps away from the root, with a standard deviation of `r sd(article_distances) %>% round(2)`. In the typical case, when an editor adds an article to an 'Australian people' subcategory, the subcategory is 2-4 steps away from the root. It seems likely that editors are mostly unaware of the category graph as a whole. Editors work locally, trying to find the best categorisations for their articles, while Wikipedia's systems work globally, enforcing the hierarchical logic of the enormous and usually invisible category graph.

If Category:Australian people can include articles about indie rock albums, then the question arises: How many of the articles are actually biographies of people connected to Australia? If we can isolate these articles, then more detailed questions become possible: How are different kinds of people represented among Wikipedia's 'Australian' biographies?

```{r}
get_persons <- function(use_cache = T) {
  all_files <- list.files(data_dir)
  these_files <- stringr::str_detect(all_files, "infobox-person-articles")
  if (!use_cache || sum(these_files) == 0) {
    # Get infobox person
    persons <- wiki_action_request() %>% 
      query_by_title("Template:Infobox person") %>% 
      query_generate_pages("transcludedin", gtilimit = "max") %>% 
      retrieve_all()
    # Get pages that transclude sub-templates of infobox person
    subtemplates <- persons %>% 
      filter(
        ns == 10,
        stringr::str_detect(title, "Infobox"),
        !stringr::str_detect(title, "/"),
        title != "Template:Infobox person")
    other_persons <- wiki_action_request() %>% 
      query_by_pageid(subtemplates$pageid) %>% 
      query_generate_pages("transcludedin", gtilimit = "max") %>% 
      retrieve_all()
    all_persons <- dplyr::bind_rows(persons, other_persons) %>% as_tibble()
    write_rds(
      all_persons,
      file.path(
        data_dir,
        glue::glue("infobox-person-articles-{as.integer(Sys.time())}.rds")
      ),
      compress = "gz"
    )
  } else {
    latest_file <- sort(all_files[these_files], decreasing = TRUE)[1]
    all_persons <- readr::read_rds(file = file.path(data_dir, latest_file))
  }
  all_persons
}
persons <- get_persons() %>% 
  filter(ns == 0) %>% 
  mutate(australian = pageid %in% australians$nodes$pageid)
```

We can identify the biographies using the [person infobox](https://en.wikipedia.org/wiki/Template:Infobox_person) and its many [subtemplates](https://en.wikipedia.org/wiki/Category:Templates_calling_Infobox_person). When an editor adds a person infobox to an article, it is a strong (though not foolproof) indicator that the article is a biography. The infobox typically appears near the start of the article, and contains structured data about the person's life, including for instance their date of birth, occupation and photograph. At the time of our analysis, `r nrow(persons) %>% prettyNum(",")` Wikipedia articles [transcluded](https://en.wikipedia.org/wiki/Help:Transclusion) a person infobox, of which `r sum(persons$australian) %>% prettyNum(",")` fell into Category:Australian people. By this metric, Australians are over-represented in English Wikipedia. They account for approximately `r round(25e6/7e9*100, 2)`% of the world's population, but `r round(sum(persons$australian) / nrow(persons) * 100, 2)`% of Wikipedia biographies.[^3]

[^3]: Of course, this `r round(sum(persons$australian) / nrow(persons) * 100, 2)`% reflects the very wide definition of 'Australianness' implicit in Category:Australian people, and may include many people who are only tangentially 'Australian'. 

*TODO*: Dig into category graph to explore gender and indigeneity.

## The search interface

TODO

## Wikidata properties

TODO

## Conclusion: beyond bias

TODO

